---
title: "Predicting Covid Data"
author: "Austin Lackey, Chien Lin and Jin Peng"
date: "Tuesday, November 6th, 2022"
output:
  powerpoint_presentation: default
  beamer_presentation: default
---

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# useful libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(pracma)
# read covid data
data <- read_csv("owid-covid-data.csv")
```

## Introduction

- In this study by Ignazio Ciufolini and Antonio Paolozzi, they use a Gauss Error function that approximates the distribution of the cumulative positive cases of COVID-19 in China
- They have then use the function to approximate the number of cumulative positive cases in Italy
- Then performed a 150 Monte Carlo simulations to predict the peak and valley of the number of daily positive cases and fatalities
- This study do not take into account factors such as daily nasopharyngeal swabs, medical, social distancing, etc. 

## The Gauss Error function

- The Gauss Error function is widely accepted as a good way to model the spread of covid data, especially in China. By using
the Gauss Error function, we can predict the spread of covid data in the future for other countries.
- $a+b*erf(c*x-d)$

## Overview of the Monte Carlo Simulation

- add more...


# China Simulation

## Available Official Data

Cumulative diagnosed positive cases:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}
# extract China data of interest
china <- data |>
  filter(location == "China") |>
  filter(date >= "2020-01-22", date <= "2020-03-28") |>
  select(c(date, total_cases, total_deaths)) |>
  mutate(days = 1:n())

# recreate plots
positive <- ggplot(data = china) +
  geom_point(aes(x = days, y = total_cases), color = "red") +
  labs(x = "Days", y = "China cumulative number of cases")
positive
```

## Available Official Data

Cumulative fatalities:

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold", out.width="50%"}
fatalities <- ggplot(data = china) +
  geom_point(aes(x = days, y = total_deaths), color = "red") +
  labs(x = "Days", y = "China cumulative number of fatalities") +
  ylim(0,5000)
fatalities
```

## Fit by the Gauss Error Function

Apply $nls$ function to determine the nonlinear (weighted) least-squares estimates of the four parameters $a,b,c,d$ of a Gauss error function model: $a+b*erf(c*x-d)$.

```{r, include=FALSE, message=FALSE, warning=FALSE}
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
```

```{r, message=FALSE, warning=FALSE}
model <- nls(total_cases ~ gauss_fn(days,a,b,c,d), 
             data=china, 
             start=list(a=10000,b=10000,c=0.1,d=1))
```

## Fit by the Gauss Error Function

Fit of cumulative diagnosed positive cases:

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
gauss_fn <- function(x,a,b,c,d) {a+b*erf(c*x-d)}
model1 <- nls(total_cases ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
summary(model1)$parameters
fn_model1 <- function(x) {4.017e+04 + 4.047e+04 * erf(9.284e-02 * x - 1.732e+00)}
positive + stat_function(fun=fn_model1, color="blue")
```

## Fit by the Gauss Error Function

Fit of cumulative fatalities:

```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
model2 <- nls(total_deaths ~ gauss_fn(days,a,b,c,d), data=china, start=list(a=10000,b=10000,c=0.1,d=1))
#summary(model2)$parameters
fn_model2 <- function(x) {1.571e+03 + 1.658e+03 * erf(6.194e-02 * x - 1.533e+00)}
fatalities +
  stat_function(fun=fn_model2, color="blue")
```

# Italy Simulation

## Plot of Diagnosed Positive Cases

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# extract Italy data
italy <- data %>% 
  filter(location=="Italy") %>% 
  filter(date <= "2020-04-21") %>%
  filter(date >= "2020-02-15") %>%
  select(date, total_cases, total_deaths) %>% 
  mutate(days = c(1:67))
italy[45:67,] <- NA


ggplot() +
  geom_point(aes(x = date, y = total_cases), data = italy)
#+ geom_smooth(aes(x = date, y = total_cases), data=italy)
```

## Plot of Fatalities

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() +
  geom_point(aes(x = date, y = total_deaths), data = italy)
#+ geom_smooth(aes(x = date, y = total_deaths), data=italy)
```

## Plot of Diagnosed Positive Cases with Gauss Error Function

```{r, echo=FALSE, message=FALSE, warning=FALSE}
x <- 1:67
ggplot(italy,aes(x)) +
  geom_point(aes(x = days, y = total_cases)) +
  stat_function(fun=function(x) 100000 + 100000 * erf(0.07*x - 3))
```

## Plot of Fatalities with Gauss Error Function

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(italy,aes(x)) +
  geom_point(aes(x = days, y = total_deaths)) +
  stat_function(fun=function(x) 11000 + 11000 * erf(0.068*x - 3))
```

# Australia Simulation

## Australia Simulation
```{r, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(pracma)
data <- read_csv("owid-covid-data.csv")
australia <- data %>% filter(location == "Australia", date > "2020-06-01", date < "2020-12-19")
australia <- subset(australia, select = c(date, total_cases, total_deaths))
australia <- australia %>% mutate(days = 1:n())
minimizedDays <- 70
australiaPre <- australia %>% filter(days < minimizedDays)
cases_Australia <- ggplot(australiaPre, aes(x = date, y = total_cases)) + 
    geom_line() + 
    labs(title = "Total Cases in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Date", y = "Total Cases")
fatalities_Australia <- ggplot(australiaPre, aes(x = date, y = total_deaths)) +
    geom_line() + 
    labs(title = "Total Deaths in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Date", y = "Total Deaths")
x <- 0:minimizedDays
cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)
cases_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_cases)) + 
    geom_line() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=2, linetype="dotted") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 1)) +
    labs(title = "Total Cases in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Cases")
fatalities_pre_plot_Australia <- ggplot(australiaPre, aes(x = days, y = total_deaths)) +
    geom_line() + 
    stat_function(fun=fatalities_function_australia, color="#3333e2", size=2, linetype="dotted") +
    scale_x_continuous(breaks = seq(0, minimizedDays, 1)) +
    labs(title = "Total Deaths in Australia from 6/1/2020 to 8/9/2020 (70 days)", x = "Days", y = "Total Deaths")
x <- 0:200
cases_function_australia <- function(x) 17200+10000*erf(0.04*x-2.4)
fatalities_function_australia <- function(x) 500+400*erf(0.038*x-3.1)
cases_plot_Australia <- ggplot(australia, aes(x = days, y = total_cases)) + 
    geom_line() + 
    stat_function(fun=cases_function_australia, color="#d33636", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 1)) +
    labs(title = "Total Cases in Australia from 6/1/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Cases")
fatalities_plot_Australia <- ggplot(australia, aes(x = days, y = total_deaths)) +
    geom_line() + 
    stat_function(fun=fatalities_function_australia, color="#3333e2", size=1.5) +
    scale_x_continuous(breaks = seq(0, 200, 1)) +
    labs(title = "Total Deaths in Australia from 6/1/2020 to 12/18/2020 (200 days)", x = "Days", y = "Total Deaths")
```
```{r, echo=FALSE}
# print(cases_Australia)
# print(fatalities_Australia)
print(cases_pre_plot_Australia)
# print(fatalities_pre_plot_Australia)
# cases_plot_Australia
# print(fatalities_plot_Australia)
```

## Australia results

```{r, echo=FALSE}
# print(cases_Australia)
# print(fatalities_Australia)
# print(cases_pre_plot_Australia)
# print(fatalities_pre_plot_Australia)
cases_plot_Australia
# print(fatalities_plot_Australia)
```

# Monte Carlo Simulations

## China

## Italy

## Australia

## Works Cited
- Link to study:
- I. Ciufolini, A. Paolozzi, Prediction of the time evolution of the Covid-19 Pandemic in Italy by a Gauss
Error Function and Monte Carlo simulations. Submitted to BioRxiv on 03.26.2020 and transferred on
03.27.2020 to MedRxiv. https://doi.org/10.1101/2020.03.27.20045104
- Link to covid data:
- https://covid19.who.int/